---
author: ""
output: 
  bookdown::pdf_document2:
    toc: false
    fig_caption: yes
    keep_tex: false
    keep_md: false
header-includes:
  - \usepackage{sectsty} %Change font sizes of section titles
  - \sectionfont{\fontsize{22}{15}\selectfont} %Change font size to 22
  - \usepackage{xcolor}
  - \definecolor{osuRed}{RGB}{215, 63, 9} %OSU red colour
  - \usepackage{setspace}
  - \usepackage{fancyhdr}
  - \usepackage{multicol}
  - \pagestyle{fancy}
  - \fancyhf{} %Remove section names from header/footer
  - \renewcommand{\headrulewidth}{0pt}
  - \usepackage{tcolorbox}
  - \usepackage{pdfpages}
  - \usepackage{tikzpagenodes}
  - \usetikzlibrary{calc}
  - \pagenumbering{gobble}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{multirow}
  - \usepackage{float}
  - \usepackage{pdflscape}
  - \usepackage{makecell} %Required for doing line breaks within a table
  - \renewcommand{\thefootnote}{\fnsymbol{footnote}} %Change footnote marks to symbols rather than numbers
  - \usepackage{tgheros} %Use TeX Gyre Heros font
  - \fontfamily{qhv}\selectfont
  - \usepackage{textcomp}
params:
  set_title: "NULL"
geometry: margin=2cm
title: ""
---

```{r setup, include=FALSE}
#OREGON BEE ATLAS PROJECT - ECOREGION REPORT SUMMARY SHEET
#WRITTEN BY SAM ROBINSON, SPRING 2026
library(knitr)
library(kableExtra)
library(ggplot2)
theme_set(theme_classic())
library(dplyr)
library(tidyr)
library(ggpubr)
library(ggrepel)
library(sf)
# library(vegan)
# library(bipartite)
library(basemaps) #https://jakob.schwalb-willmann.de/basemaps/ #Open source mapping package
library(stars)

#knitr options
knitr::opts_chunk$set(
	eval = TRUE,
	echo = FALSE,
	fig.path = "./figs",
	message = FALSE,
	warning = FALSE
)

options(dplyr.summarise.inform=FALSE, warn=-1) #Suppress warnings
  
if(any(sapply(params,function(x) x=='NULL'))){
  stop("Vineyard name and output directory must be specified")
}

#Load vineyard plant data from iNaturalist records

#Plants from focal project
iNatPlDat_vy <- filter(iNatPlDat, vineyard==params$set_title)

#Debugging: param is a reserved word in this context

iNatPlDat_vy <- filter(iNatPlDat, vineyard=='callotchrissy')

if(length(unique(iNatPlDat_vy$ecoreg))!=1){
  stop(paste0('Project has ',length(unique(iNatPlDat_vy$ecoreg)),' ecoregions; should be 1'))
} else {
  ecoRegName <- unique(iNatPlDat_vy$ecoreg)
}

isMultiyear <- length(unique(iNatPlDat_vy$year))>1 #More than 1 year of data?

#Ecoregion-specific bee data
beeData_ecoreg <- beeData %>% filter(ecoreg==ecoRegName) %>% 
  st_drop_geometry() %>% 
  select(-CollectorName,-Sex,-Method,-ecoreg)

#Plant genera found on project during different years
inatPlDat_year <- iNatPlDat_vy %>% st_drop_geometry() %>% 
  select(plGenus,year) %>% distinct() %>% 
  mutate(across(everything(),factor)) %>% 
  count(plGenus,year,.drop = FALSE) %>% 
  pivot_wider(names_from=year,values_from = 'n',names_prefix = 'in',
              values_fill = 0,values_fn = ~.x==1)
# inatPlDatSpp_year <- iNatPlDat_vy %>% st_drop_geometry() %>% #Same as above, but for full scientific name
#   select(scientific_name,year) %>% distinct() %>% filter(!grepl('spp.',scientific_name)) %>% 
#   count(scientific_name,year) %>% pivot_wider(names_from=year,values_from = 'n',names_prefix = 'in',
#               values_fill = 0,values_fn = ~.x==1)

#Bee records from region that are found on plant genera from iNat project
# dat_inatPlant <- beeData_ecoreg %>% filter(plantGenus %in% unique(iNatPlDat_vy$plGenus))
# datSpp_inatPlant <- beeData_ecoreg %>% filter(foragePlant %in% unique(iNatPlDat_vy$scientific_name)) #Same as above, but for full scientific name
dat_inatPlant <- beeData_ecoreg %>% st_drop_geometry() %>% 
  right_join(inatPlDat_year,by=c('plantGenus'='plGenus')) #Joins in yearly occurrence of each plant genus

#Get scores for potential new plant genera/species

#Scores for potential new spp/genera
plantScoresSpp <- getPlGenScores(mat = ecoRegNetworks[[ecoRegName]]$ntwk_all, #Regional network (plant spp - bee spp)
                               plGen = unique(iNatPlDat_vy$scientific_name[!grepl('spp.',iNatPlDat_vy$scientific_name)]), #Plants from focal vineyard
                               exclude = nonNativeSpp[!grepl('spp.',nonNativeSpp)]) %>% #Non-native species to exclude
  select(-Chao1,-Chao1Z) %>% 
  left_join(plantList,by=c('PlantGenus'='Scientific_name')) %>% 
  filter(!is.na(Common_name)) #Removes genSpp not found in plant list - mostly errors from the recorded plants in the bee data, I think

#Genus-genus interaction matrix only for iNat plant genera
inatGenMat <- ecoRegNetworks[[ecoRegName]]$ntwk_gen_all[rownames(ecoRegNetworks[[ecoRegName]]$ntwk_gen_all) %in% unique(iNatPlDat_vy$plGenus),] 

#iNat plant genera that have no visiting bees (using data from the regional network)
badPlGen <- sort(unique(iNatPlDat_vy$plGenus)[!unique(iNatPlDat_vy$plGenus) %in% unique(beeData_ecoreg$plantGenus)])

#Regional bee genera and species that are supported by listed plant genera
# suppBeeGen <- dat_inatPlant %>% filter(!is.na(Species)) %>% pull(Genus) %>% unique() #Supported
# suppBeeSpp <- dat_inatPlant %>% filter(!is.na(Species)) %>% pull(genSpp) %>% unique()

#Predicted (yearly) bee occurrance based on genera found on the project (bee "alpha diversity")
suppBeeGen <- dat_inatPlant %>% filter(!is.na(Genus)) %>% 
  select(Genus,contains('in20')) %>% distinct() %>% #Bee genera 
  group_by(Genus) %>% summarize(across(contains('in20'),~any(.x))) #%>% 
  # left_join(famGen,by='Genus') 

suppBeeSpp <- dat_inatPlant %>% filter(!is.na(Species)) %>% 
  select(genSpp,contains('in20')) %>% distinct() %>% #Bee species
  group_by(genSpp) %>% summarize(across(contains('in20'),~any(.x))) %>% 
  mutate(Genus=gsub('\\s.*','',genSpp)) #%>% 
  # left_join(famGen,by='Genus') %>% select(-Genus)

#Regional bee diversity (bee "gamma diversity")
regBeeGen <- beeData_ecoreg %>% filter(!is.na(Species)) %>% select(Genus) %>% distinct() #Bee genera
regBeeSpp <- beeData_ecoreg %>% filter(!is.na(Species)) %>% select(genSpp) %>% distinct() #Bee species

#Counts of supported bees by family (existing and suggested 5 new plants)
famCounts <- bind_rows(mutate(suppBeeSpp,cat='existing'),
                       data.frame(genSpp=trimws(strsplit(paste0(plantScoresSpp$newSpp[1:5],collapse=', '),split = ',')[[1]]),
                                  cat='new')) %>% 
  mutate(Genus=gsub('\\s.+$','',genSpp)) %>% left_join(famGen,by='Genus')

suppBeeGen <- suppBeeGen %>% left_join(famGen,by='Genus')
suppBeeSpp <- suppBeeSpp %>% left_join(famGen,by='Genus') %>% select(-Genus)

#Estimate regional bee species richness and abundance for each plant species 
rareBees <- ecoRegNetworks[[ecoRegName]]$rareBees

#Regional plant species visitor richness + other traits - includes non-visited plants
plSppRich <- ecoRegNetworks[[ecoRegName]]$topSpp %>% #Visited plants from ecoregion
  full_join(plantList,by=c('plantSpp'='Scientific_name')) %>% #Join to plant spp data
  filter(!grepl('spp.',plantSpp)) %>% #Remove genus-level records
  transmute(genSpp=plantSpp,N=Nbees,EstRich=Nbees_estim,Nrare=Nbees_rare, #Choose fields of interest
            Synonym,Common_name,isWeedy,isNoxious,isNative,
            Lifecycle,Bloom_start,Bloom_end,quality) %>% 
  mutate(across(where(is.factor),as.character)) %>% #Factor to character
  mutate(across(c(isNative,isNoxious),~ifelse(is.na(.x),FALSE,.x))) %>% #Sets unknown origin plants to exotic & non-noxious
  mutate(across(c(N,EstRich,Nrare),~ifelse(is.na(.x),0,.x)), #Fills in unmatched records with zeros (not visited)
         quality=ifelse(is.na(quality),'poor',quality)) 

#Regional plant genus visitor richness + other traits
plGenRich <- ecoRegNetworks[[ecoRegName]]$topGen %>% mutate(plantGen = paste0(plantGen,' spp.')) %>% 
  full_join(filter(plantList,grepl('spp.',Scientific_name)),by=c('plantGen'='Scientific_name')) %>% 
  transmute(genSpp=plantGen ,N=Nbees,EstRich=Nbees_estim,Nrare=Nbees_rare,
            Synonym,Common_name,isWeedy,isNoxious,isNative,
            Lifecycle,Bloom_start,Bloom_end,quality) %>% 
  mutate(genSpp=gsub(' spp.','',genSpp)) %>% 
  mutate(across(where(is.factor),as.character)) %>% #Factor to character
  mutate(across(c(isNative,isNoxious),~ifelse(is.na(.x),FALSE,.x))) %>% #Sets unknown origin plants to exotic & non-noxious
  mutate(across(c(N,EstRich,Nrare),~ifelse(is.na(.x),0,.x)),#Fills in unmatched records with zeros (not visited)
         quality=ifelse(is.na(quality),'poor',quality)) %>% 
  mutate(quality=factor(quality,levels=c('poor','good','super')))

#Join regional info to (grouped) grower plant info
inatPlDat_spp <- iNatPlDat_vy %>% st_drop_geometry() %>% 
  filter(!grepl('.spp',scientific_name)) %>% #Removes genus only records
  group_by(scientific_name) %>% summarize(common_name=first(common_name),Nplants=n()) %>% #N = Number of bee spp recorded
  left_join(plSppRich,c('scientific_name'='genSpp')) %>% 
  mutate(Common_name=ifelse(is.na(Common_name),common_name,Common_name)) %>% select(-common_name) %>% #Replaces common name from plant list with iNat name if NA
  mutate(across(where(is.factor),as.character)) %>% #Factor to character
  mutate(across(c(N,EstRich,Nrare),~ifelse(is.na(.x),0,.x)),#Fills in unmatched records with zeros (not visited)
         quality=ifelse(is.na(quality),'poor',quality)) %>% 
  mutate(quality=factor(quality,levels=c('poor','good','super')))

inatPlDat_gen <- iNatPlDat_vy %>% st_drop_geometry() %>% 
  group_by(plGenus) %>% summarize(common_name=first(common_name),Nplants=n()) %>% 
  left_join(plGenRich,by=c('plGenus'='genSpp')) %>% 
  mutate(Common_name=ifelse(is.na(Common_name),common_name,Common_name)) %>% select(-common_name) %>% #Replaces common name from plant list with iNat name if NA
  mutate(across(where(is.factor),as.character)) %>% #Factor to character
  mutate(across(c(N,EstRich,Nrare),~ifelse(is.na(.x),0,.x)),#Fills in unmatched records with zeros (not visited)
         quality=ifelse(is.na(quality),'poor',quality)) %>% 
  mutate(quality=factor(quality,levels=c('poor','good','super')))

#"Super bee plants" - must be high-quality and not noxious
superPl_spp <- inatPlDat_spp %>% 
  filter(!is.na(N),quality=='super',!isNoxious) %>% 
  arrange(desc(N),desc(Nrare))

superPl_gen <- inatPlDat_gen %>% 
  filter(!is.na(N),quality=='super',!isNoxious) %>% 
  arrange(desc(N),desc(Nrare))

#Get range of bees found on other projects from this ecoregion
beeDivRange <- sapply(iNatNetworks,function(x,y) ifelse(x$ecoregNtwk_summary$ecoRegName!=y,NA,ncol(x$ntwk)),
       y=ecoRegName)
beeDivRange <- unique(unname(beeDivRange[!is.na(beeDivRange)])) 

# \node[draw=black,minimum height = 2.3cm,minimum width=8cm,anchor=north east,font=\color{white}\LARGE\linespread{1}\selectfont, fill=\color{white},rectangle,yshift=-0.5cm,xshift=-0.5cm,align=right] at (current page.north east){BEE HABITAT\\Stewardship Tool};


    # <!-- \node[anchor=north west,rectangle,,yshift=0cm,xshift=0.5cm] at (current page.north west){\includegraphics[height=2.3cm]{../extdata/OSU_ES-MM_horizontal_1C_W.png}}; -->
    # <!-- \node[anchor=south east,rectangle,,yshift=0.2cm,xshift=-4.5cm] at (current page marginpar area.north east){\includegraphics[width=4.5cm]{../extdata/OregonBeeAtlas-Logo-Final-03.png}}; -->

```

\newgeometry{top=2in}

<!-- Create overlay for beginning of document-->

\begin{tikzpicture}[remember picture,overlay]
    \fill[osuRed] (current page.north west) rectangle ([yshift=0cm] current page text area.north  -| current page.east); 
    \node[minimum height = 2.3cm,minimum width=8cm,anchor=south west,font=\color{white}\Huge\linespread{1}\selectfont,rectangle,yshift=-3.5cm,xshift=2.1cm,align=left] at (current page.north west){\textsc{Stewardship Report}};
    \node[anchor=south east,rectangle,,yshift=0.2cm,xshift=0cm] at (current page marginpar area.north east){\includegraphics[height=4cm]{../extdata/blackwhite_beestewards_logo.png}};
\end{tikzpicture}

\vspace{0.2cm}

\centering\huge

\huge 2025 Report for `r params$set_title` 

\raggedright\Large

\section{Plant inventory}

\setlength{\columnsep}{-8.5cm}

\begin{multicols}{2}[
Out of the `r nrow(inatPlDat_spp)` identified plant species from your iNaturalist project, \Large`r nrow(filter(inatPlDat_spp,!is.na(N)&quality!='poor'))` \Large are bee-friendly plants. Of these plants:
]

\includegraphics[width=2.5cm]{../extdata/purple_flowers.png}

\columnbreak

\begin{itemize}
\item \huge`r nrow(superPl_spp)` \Large species are \emph{super} bee plants
\item \huge`r nrow(filter(inatPlDat_spp,!is.na(N)&quality!='poor'&isNative))` \Large species are native to your Ecoregion
\item \huge`r nrow(filter(inatPlDat_spp,!is.na(N)&quality!='poor'&isWeedy))` \Large species are not native to your Ecoregion
\item \huge`r nrow(filter(inatPlDat_spp,!is.na(N)&quality!='poor'&isNoxious))` \Large species are noxious weeds
\end{itemize}

\end{multicols}

<!-- This gets the number of bee species predicted from plant genera (not plant species)-->

\section{Bee inventory}

Based on bee-plant records from Oregon Bee Atlas, we predict that you have \Large`r nrow(suppBeeSpp)` \Large species of bees at your site. `r ifelse(length(beeDivRange)>1,paste0('Similar sites in the ',params$set_title,' ecroegion have ',paste0(range(beeDivRange),collapse=' - '),' species of bees.'),'')`
The species at your site cover `r length(unique(filter(famCounts,cat=='existing')$lookupFam))` out of the 6 bee families found in North America.

```{r beeFamilyPlot, out.width = '100%', fig.asp=0.35}

if(isMultiyear){
  #Shows additional spp added each year
  famCounts %>% filter(cat=='existing') %>%
    mutate(lookupFam=factor(lookupFam,levels= rev(commonFam$Family),labels = rev(commonFam$plotLab))) %>% 
    pivot_longer(cols=contains('in20'),names_to='year',names_transform = as.factor) %>%
    group_by(genSpp) %>% mutate(value=cumsum(value)>0) %>% ungroup() %>% filter(value) %>%
    count(lookupFam,year,.drop=FALSE) %>% pivot_wider(names_from='year',values_from='n') %>% 
    pivot_longer(cols=contains('in20'),values_to='n',names_to='year') %>% 
    group_by(lookupFam) %>% mutate(n=ifelse(is.na(n-lag(n)),n,n-lag(n))) %>% ungroup() %>% 
    filter(n>0) %>% 
    mutate(year=factor(year,levels=rev(unique(year)),labels=gsub('in','',rev(unique(year))))) %>%
    ggplot(aes(x=lookupFam,y=n))+
    geom_col(aes(fill=year),col='black')+
    labs(x=NULL,y='Number of species',fill='Year')+
    coord_flip()+
    scale_fill_brewer(palette = 'Greys',direction = -1)
  
} else {
  famCounts %>% filter(cat=='existing') %>%
    mutate(lookupFam=factor(lookupFam,levels= rev(commonFam$Family),labels = rev(commonFam$plotLab))) %>%
    count(lookupFam,.drop=FALSE) %>%
    ggplot(aes(x=lookupFam,y=n))+
    geom_col(col='black',fill='grey20')+
    labs(x=NULL,y='Number of species')+
    coord_flip()
}

```


\restoregeometry

\clearpage

\section{Glossary}

\begin{table}[h]
\Large
\begin{tabular}{p{5cm}p{12cm}}
\textbf{Ecoregion}       & An ecological region made of ecosystems that   have similar climate, geology, soil, and ecology              \\ [5ex]
\textbf{Super bee plant} & A plant species that attracts a wide diversity of bees, rare specialist   bees, and is native to your region \\ [5ex]
\textbf{Exotic plant}    & A plant species that is not native or naturalized in your region                                             \\ [5ex]
\textbf{Noxious weed}    & A plant species that is not native or naturalized in your region, that   causes economic or ecological harm 
\end{tabular}
\end{table}


\section{The \emph{super bee plants} at your site}

`r if(nrow(superPl_spp)>0) 'These species are doing the work at your site to support a diverse community of bees.'`
`r if(nrow(superPl_spp)<=0) "We didn’t find (m)any super bee plants at your site. This might be because they weren't documented in your iNaturalist project, or because your site has a lot of opportunity for growth! On the next page, we've put together a site-specific list of plant species to take your bee habitat up a notch and expand the list of bees your site can support."`

```{r superBeePlantsSpp}
numPlants <- 8 #Max number of plants to display

if(nrow(superPl_spp)>0){
  superPl_spp %>%
  select(Common_name,scientific_name,EstRich,Nrare,Lifecycle,Bloom_start,Bloom_end) %>%
  slice(1:numPlants) %>%
  transmute(Plant=linebreak(paste0(Common_name,'\n(\\emph{',scientific_name,'})'),align='r'),
            `Life cycle`=Lifecycle,
            `Bloom Time`=paste0(Bloom_start,'-',Bloom_end)) %>%
  mutate(across(where(is.character),~ifelse(grepl('NA',.x),'-',.x))) %>% #Set NAs as dashes
  kable(escape=FALSE,format = 'latex',
        position = "h",longtable=TRUE) %>%
    kable_styling(latex_options = c("repeat_header")) #%>% 
    # column_spec(2,width='3cm') 
} else {
  
}
```

`r if(nrow(superPl_spp)>0) 'Super bee plants on your property (in order of importance). This data is site-specific, which does not allow for comparisons with other sites.'`

\Large
\clearpage
\section{Where are the \emph{super bee plants} located?}

Bee plants are often located in overlooked areas. While some plants are in your cultivated landscaped areas, often they are found growing wild in your headlands. Consider increasing the number of plants in these areas through seeding or cuttings, or by managing weeds and grasses.

```{r fieldMapsSpp, fig.asp=0.75, fig.width=7, warning=FALSE}

plantLocs <- iNatPlDat_vy %>% select(-vineyard,-ecoreg) %>% 
  # st_transform(4269) %>% mutate(lon=st_coordinates(.)[,1],lat=st_coordinates(.)[,2]) %>% st_drop_geometry() %>%
  left_join(select(plSppRich,genSpp,isNoxious,quality),by=c('scientific_name'='genSpp')) %>%
  left_join(select(plGenRich,genSpp,isNoxious,quality),by=c('plGenus'='genSpp')) %>%
  mutate(across(contains('quality'),as.character)) %>% 
  
  mutate(isNoxious.y=ifelse(!grepl('spp.',scientific_name),NA,isNoxious.y),
         quality.y=ifelse(!grepl('spp.',scientific_name),NA,quality.y)) %>%
  mutate(isNoxious.x=ifelse(is.na(isNoxious.x),isNoxious.y,isNoxious.x),
         quality.x=ifelse(is.na(quality.x),quality.y,quality.x)) %>%
  
  select(-isNoxious.y,-quality.y) %>% mutate(quality.x=ifelse(is.na(quality.x),'poor',quality.x)) %>%
  mutate(quality.x=factor(quality.x,levels=c('poor','good','super')),
         taxonLevel=ifelse(grepl('.spp',scientific_name),'Genus','Species')) %>% 
   st_jitter(amount = 3) %>% arrange(quality.x) #Sorts, so super plants are rendered on top

# ext <- st_as_sfc(st_bbox(plantLocs), crs=st_crs(plantLocs)) %>% st_transform(3857)
# b <- basemap_stars(ext, map_service = "esri", map_type = "world_imagery",verbose = FALSE)
# 
# #Species-level map
# ggplot() + geom_stars(data=st_rgb(b))+
#   geom_sf(data=st_transform(filter(plantLocs,taxonLevel=='Species'),st_crs(b)),
#           aes(col=quality.x),show.legend=TRUE)+
#   geom_text_repel(data = filter(st_transform(plantLocs,st_crs(b)),taxonLevel=='Species',quality.x=='super'),
#     aes(label = common_name, geometry = geometry),
#     stat = "sf_coordinates",col='cyan',size=3)+
#   labs(x=NULL,y=NULL,col='Plant quality',title='Species-level IDs')+
#   theme(axis.text = element_blank(),axis.ticks = element_blank())+
#   scale_colour_manual(values=c('tomato3','purple','cyan'),drop=FALSE)

```

\normalsize
\vspace{0.5cm}

\Large


\clearpage
\section{Plant recommendations}

There are a few native bee plants from the Willamette Valley region you are currently missing. Adding these species of plants will increase the estimated number of bees at your property from \Large`r as.character(sum(famCounts$cat=='existing'))` to \Large `r nrow(famCounts)` species of bees: \large 

\vspace{0.5cm}

```{r plantsToAdd}

addRotate <- function(x) paste0('\\rotatebox[origin=c]{90}{',x,'}')

nPlantsMax <- 5 #Max number of suggested plants (annuals/perennials)

data.frame(genSpp=plantScoresSpp$PlantGenus) %>% 
  left_join(select(plantList,Common_name,Scientific_name,isNative,Lifecycle,Bloom_start,Bloom_end,isNative:isWetland),
            by=c('genSpp'='Scientific_name')) %>% 
  mutate(Rank=1:n(),Plant=linebreak(paste0(Common_name,'\n(\\emph{',genSpp,'})'),align='r'),
            Origin=ifelse(isNative,'Native','Exotic'),
            Bloom_period=paste0(Bloom_start,'-',Bloom_end)) %>%
  select(Rank,Plant,Lifecycle,Bloom_period) %>% 
  rename(`Bloom Period`=Bloom_period,`Plant Species`=Plant) %>% 
  group_by(Lifecycle) %>% arrange(desc(Rank)) %>% slice(1:nPlantsMax) %>% ungroup() %>% select(-Rank) %>% 
  kable(escape=FALSE,format = 'latex',longtable=TRUE) %>% 
  kable_styling(latex_options = c("repeat_header","scale_down")) %>% 
  column_spec(2:3,width='2cm',latex_valign = 'm') #Warning: argument not atomic vector
  

```

For a full list of native plants, see Appendix 3.

\newpage

```{r beeFamilyPlotNew, out.width = '100%', fig.asp=0.4}
famCounts %>% 
  mutate(lookupFam=factor(lookupFam,levels= rev(commonFam$Family),labels = rev(commonFam$plotLab))) %>%
  mutate(cat=factor(cat,levels=c('new','existing'),labels=c('Future','Current'))) %>%
  count(lookupFam,cat,.drop=FALSE) %>%
  ggplot(aes(x=lookupFam))+
  geom_col(aes(y=n,fill=cat),col='black',show.legend = TRUE)+
  labs(x=NULL,y='Number of species',fill=NULL,title='Target for bee conservation')+
  scale_fill_manual(values=c('darkorange','grey20'))+
  coord_flip()
  
```

\Large
\vspace{0.3cm}

\subsection{Note on implementation}

\textbf{Soil and Water Conservation Districts (SWCD)} offer free technical assistance and site visits to County residents interested in implementing conservation projects on their land. They can help develop a pollinator planting plan that is specific to your site and operation. In addition to pollinator plantings, landowners with streams and riparian areas may be eligible to enroll in a financial assistance program, the \textbf{Conservation Reserve Enhancement Program (CREP)}, to establish native trees and shrubs along creeks. Reach out to your SWCD with this report to help you move forward with restoration plans.

\clearpage

\section*{Appendix 1: Predicted bees from your property}

```{r allBees}
suppBeeSpp %>% select(-contains('in')) %>% rename(Family=lookupFam) %>%
  group_by(Family) %>% summarize(`Bee Species`=paste0(genSpp,collapse=', '),N=n()) %>%
  ungroup() %>% mutate(Family=paste0(Family,'\n(',N,' species)')) %>% select(-N) %>%
  kable(escape=FALSE,format = 'latex',align = c('r','l'),longtable=TRUE) %>%
  # kable_styling(full_width = TRUE) %>% 
  column_spec(1,bold = TRUE,width='5cm') %>%
  column_spec(2,width='10cm',italic = TRUE)
```

\clearpage

\begin{landscape}

\section*{Appendix 2: Full super-bee plant list}

```{r fullPlantListSpp}
if(nrow(superPl_spp)>0){
  superPl_spp %>%
  select(Common_name,scientific_name,EstRich,Nrare,isNative,Lifecycle,Bloom_start,Bloom_end) %>%
  transmute(Plant=linebreak(paste0(Common_name,'\n(\\emph{',scientific_name,'})'),align='r'),
            `Life cycle`=Lifecycle,
            `Bloom Time`=paste0(Bloom_start,'-',Bloom_end),
            `Bee Species`=round(EstRich),`Rare Bees`=Nrare) %>%
  mutate(across(where(is.character),~ifelse(grepl('NA',.x),'-',.x))) %>% #Set NAs as dashes
  kable(escape=FALSE,format = 'latex',caption = 'Super bee plant species',longtable=TRUE) %>%
  # kable_styling(full_width = TRUE) %>% 
  column_spec(2,width='3cm')
}
```

\end{landscape}

